Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-08-01 Moltres.tex
extraneous 'that'

Moltres is an open source code licensed under \gls{LGPL} terms so that the 
\gls{MSR} community can freely use, interrogate, and improve it.  Its openness 
is a defining characteristic and promotes
quality through transparency and ease of peer review. In the era of
GitHub\cite{noauthor_build_nodate} and international scientific collaboration,
open and modern software practices must be employed in order for nuclear
engineering simulation capability to enable discovery and support the regulatory
needs presented by new reactor designs. In that vein, Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to
enable extension and code reuse. Moreover, Moltres depends on the \gls{MOOSE} 
framework, \cite{gaston_physics-based_2015} another \gls{LGPL} code that itself 
leans on LibMesh \cite{kirk_libmesh:_2006}, a
\gls{LGPL} finite element library, and PetSc \cite{satish_balay_petsc_2015}, a
\gls{BSD}-licensed toolkit for solving nonlinear equations yielded by 
discretizing PDEs. \gls{MOOSE} and LibMesh translate weak PDE forms defined by
applications (e.g. Moltres) into residual and Jacobian
functions. These functions are the inputs into PetSc Newton-Raphson solution routines. All
codes use MPI for parallel communication and are easily deployed on
massively-parallel cluster-computing platforms. \gls{MOOSE} applications by
default use monolithic and implicit methods that are ideal for closely-coupled
and multi-scale physics, such as the model problem described
in this work. However, Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it 
extensible to myriad future modeling challenges.

@@

Moltres \cite{lindsay_arfc/moltres_nodate} is implemented as an application for
use atop the \gls{MOOSE} \cite{gaston_physics-based_2015} framework.
Accordingly, Moltres includes physics kernels and boundary conditions for
solving for neutron fluxes, temperature, and precursor concentrations. In \gls{MOOSE}
jargon, kernels are C++ classes that contain methods for computing residual and
Jacobian contributions corresponding to individual pieces of governing
equations. Developing the code-base in this way allows modular construction
of equation systems; e.g. the kernel used to represent heat
conduction can also represent generic chemical diffusion. Moltres
also features neutron and precursor ``actions.'' These actions automatically 
construct the systems of equations for an arbitrary number of neutron and
precursor groups. This means that as long as group constants are provided in an appropriate
tabular form, a user only has to modify a couple of lines in a Moltres input
file to change from say two to forty-four neutron groups.

@@

The model molten salt reactor closely emulates the \gls{MSRE}. When developing new physics or investigating
different types of transients, one can reduce the model problem to a two
dimensional axisymmetric one for rapid proof of concept. To approximately simulate the lattice
structure of the \gls{MSRE} under 2D conditions, a geometry is constructed with 14 repeating
fuel-moderator regions, as shown in \Cref{fig:geom}. The fuel and moderator
radii are chosen such that the resulting area/volume fraction of fuel is .225 as
for the \gls{MSRE}.

@@

\Cref{fig:axial_fluxes_compare} compares the axial flux profiles calculated by
Moltres and the \gls{ORNL} \gls{MSRE} design model. Note that the radius for the
plots are chosen to correspond to the peak of the thermal flux in both cases;
for the \gls{ORNL} calculations this is 8.4 inches from the core center-line
because of the effect of the control rod thimbles and extra fuel along the
center-line. Once again, the plots are in decent agreement. The \gls{ORNL}
calculations include the lower and upper plena which are not included in this
report's Moltres model. Consequently, the plot in \cref{fig:msre_axial_fluxes}
extends to lower and higher z-values than
\cref{fig:moltres_axial_fluxes}. Additionally, absorption in the plena caused 
deviation of the thermal flux from a sinusoidal shape in 
\cref{fig:msre_axial_fluxes}.

>>>>

Moltres is an open source code licensed under \gls{LGPL} terms so the 
\gls{MSR} community can freely use, interrogate, and improve it.  Its openness 
is a defining characteristic and promotes
quality through transparency and ease of peer review. In the era of
GitHub\cite{noauthor_build_nodate} and international scientific collaboration,
open and modern software practices must be employed in order for nuclear
engineering simulation capability to enable discovery and support the regulatory
needs presented by new reactor designs. In that vein, Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to
enable extension and code reuse. Moreover, Moltres depends on the \gls{MOOSE} 
framework, \cite{gaston_physics-based_2015} another \gls{LGPL} code that itself 
leans on LibMesh \cite{kirk_libmesh:_2006}, a
\gls{LGPL} finite element library, and PetSc \cite{satish_balay_petsc_2015}, a
\gls{BSD}-licensed toolkit for solving nonlinear equations yielded by 
discretizing PDEs. \gls{MOOSE} and LibMesh translate weak PDE forms defined by
applications (e.g. Moltres) into residual and Jacobian
functions. These functions are the inputs into PetSc Newton-Raphson solution routines. All
codes use MPI for parallel communication and are easily deployed on
massively-parallel cluster-computing platforms. \gls{MOOSE} applications by
default use monolithic and implicit methods ideal for closely-coupled
and multi-scale physics, such as the model problem described
in this work. However, Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it 
extensible to myriad future modeling challenges.

@@

Moltres \cite{lindsay_arfc/moltres_nodate} is implemented as an application for
use atop the \gls{MOOSE} \cite{gaston_physics-based_2015} framework.
Accordingly, Moltres includes physics kernels and boundary conditions for
solving for neutron fluxes, temperature, and precursor concentrations. In \gls{MOOSE}
jargon, kernels are C++ classes that contain methods for computing residual and
Jacobian contributions corresponding to individual pieces of governing
equations. Developing the code-base in this way allows modular construction
of equation systems; e.g. the kernel used to represent heat
conduction can also represent generic chemical diffusion. Moltres
also features neutron and precursor ``actions.'' These actions automatically 
construct the systems of equations for an arbitrary number of neutron and
precursor groups. Therefore, as long as group constants are provided in an appropriate
tabular form, a user only has to modify a couple of lines in a Moltres input
file to change from say two to forty-four neutron groups.

@@

The model molten salt reactor closely emulates the \gls{MSRE}. When developing new physics or investigating
different types of transients, one can reduce the model problem to a two
dimensional axisymmetric one for rapid proof of concept. To approximately simulate the lattice
structure of the \gls{MSRE} under 2D conditions, a geometry is constructed with 14 repeating
fuel-moderator regions, as shown in \Cref{fig:geom}. The fuel and moderator
radii are chosen such that the resulting area/volume fraction of fuel is 0.225 as
for the \gls{MSRE}.

@@

\Cref{fig:axial_fluxes_compare} compares the axial flux profiles calculated by
Moltres and the \gls{ORNL} \gls{MSRE} design model. The radius for the
plots are chosen to correspond to the peak of the thermal flux in both cases;
for the \gls{ORNL} calculations this is 8.4 inches from the core center-line
because of the effect of the control rod thimbles and extra fuel along the
center-line. Once again, the plots are in decent agreement. The \gls{ORNL}
calculations include the lower and upper plena which are not included in this
report's Moltres model. Consequently, the plot in \cref{fig:msre_axial_fluxes}
extends to lower and higher z-values than
\cref{fig:moltres_axial_fluxes}. Additionally, absorption in the plena caused 
deviation of the thermal flux from a sinusoidal shape in 
\cref{fig:msre_axial_fluxes}.
